Understanding the Impact of Marital Status, Income, and Age on Individuals' Poverty Status

### 2024-03-29 paper/paper.qmd
mortgage and income

```{r}

```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

@@

@fig-post_prior compares the posterior with the prior. This compares the prior distribution of parameters with the posterior distribution of parameters in our logistic regression model. We can see that half of the model parameters do not change after data are taken into account, while some parameters shift slightly. This shows that the observed data partially matches with our initial belief and expectation about the poverty status of people in the United States in 2019. We can see that for people with income level at "below 10k", the posterior distribution greatly shifts from its prior after we input observed data. This is likely suggesting that the observed data for "below 10k" strongly contradicts our initial belief. This suggests that the majority of people in US who earns less than a total amount of 10k per household annually is in poverty in 2019.

@@

# Convert variables to factors
analysis_data$mortgage <- factor(analysis_data$mortgage)
analysis_data$income <- factor(analysis_data$income)
analysis_data$age <- factor(analysis_data$age)

@@

posterior_vs_prior(poverty_prediction_model) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and mortgage"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of owner without mortgage", "Trace plot of renter"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "(Intercept)")
plot(poverty_prediction_model, "trace", "mortgageOwner without mortgage")
plot(poverty_prediction_model, "trace", "mortgageRenter")
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace3
#| fig-cap: "Trace plot of employ"
#| fig-subcap: ["Trace plot of age 35-44", "Trace plot of age 45-54", "Trace plot of age 55-64", "Trace plot of age above 65"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "age35-44")
plot(poverty_prediction_model, "trace", "age45-54")
plot(poverty_prediction_model, "trace", "age55-64")
plot(poverty_prediction_model, "trace", "ageabove 65")
```

@@

plot(poverty_prediction_model, "rhat")
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden with x_axis limits"
# Create the model plot
model_plot <- modelplot(poverty_prediction_model, conf_level = 0.90, size = 0.2)

# Modify the x-axis limits
model_plot + xlim(-20, 20) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```

>>>>

## Data Source

## Features

# read in raw data
raw_poverty_data <-  read_csv(here::here("data/raw_data/raw_poverty_data.csv"))
```

```{r}
#| label: tbl-raw
#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

raw_poverty_data[1:5, 1:4] |>
  kable()
```

## Data Measurement

## Methodology

# read in cleaned data
poverty2019 <-  read_parquet(here::here("data/analysis_data/cleaned_poverty_data.parquet"))

poverty2019 <-
  poverty2019 |>
  mutate(
    poverty_status = as_factor(poverty_status),
    mortgage_state= factor(
      mortgage_state,
      levels = c(
      "Owner with Mortgage",
      "Owner without Mortgage or rent-free",
      "Renter")
    ),
    income = factor(
      income,
      levels = c(
      "below 10k",
      "10k-50k",
      "50k-100k",
      "100k-150k",
      "150k-200k",
      "200k-250k",
      "above 250k")
    )
  ) |>
  select(poverty_status, mortgage_state, income)
```

```{r}
#| label: tbl-clean-data
#| tbl-cap: "Preview of the cleaned 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

poverty2019[5:9, 1:4] |>
  kable()
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

poverty2019 |>
  summary() |>
  kable(align = "c") %>%
  kable_styling(font_size = 8)
```

## Data Visualization
```{r}
#| label: fig-mortgage
#| fig-cap: The distribution of poverty status by mortgage state
#| echo: false
#| warning: false
#| message: false
poverty2019 |>
  ggplot(aes(x = mortgage_state, fill = poverty_status)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=6)) +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Mortgage state",
    y = "Number of respondents",
    fill = "Poverty status"
  )
```
```{r}
#| label: fig-income
#| fig-cap: The distribution of poverty status by income level
#| echo: false
#| warning: false
#| message: false
poverty2019 |>
  ggplot(aes(x = income, fill = poverty_status)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Income level",
    y = "Number of respondents",
    fill = "Poverty status"
  )
```

@@

@fig-post_prior compares the posterior with the prior. This compares the prior distribution of parameters with the posterior distribution of parameters in our logistic regression model. We can see that half of the model parameters do not change after data are taken into account, while some parameters shift slightly. This shows that the observed data partially matches with our initial belief and expectation about the poverty status of people in the United States in 2019. We can see that for people with income level at "below 10k" and "10k-50k", the posterior distributions shift from their prior after we input observed data; their distribution not crossing 0 at all. This is suggesting that the observed data for "below 10k" and "10k-50k" strongly contradict our initial belief. So the majority of people in US who earns less than a total amount of 50k per household annually was in poverty in 2019. Also, the intercept (people who are owners with mortgage and earn 100-150k) shifts to the left; its distribution not crossing 0. This indicates that the actual observation does not match with our prior belief. People in this category was less likely to be in poverty state in 2019 United States.

@@

# Convert variables to factors
analysis_data$marital_status <- factor(analysis_data$mortgage_state)
analysis_data$income <- factor(analysis_data$income)

@@

posterior_vs_prior(poverty_prediction_model, size = 0.1) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=8)) +
  guides(color = guide_legend(ncol = 4))+
  coord_flip()
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and marital status"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Owner without Mortgage or rent-free", "Trace plot of Renter"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "(Intercept)")
plot(poverty_prediction_model, "trace", "mortgage_stateOwner without Mortgage or rent-free")
plot(poverty_prediction_model, "trace", "mortgage_stateRenter")
```

@@

@@

plot(poverty_prediction_model, "rhat")
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of positive poverty status"

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of positive poverty status with x_axis limits"

# Create the model plot
model_plot <- modelplot(poverty_prediction_model, conf_level = 0.90, size = 0.2)

# Modify the x-axis limits
model_plot + xlim(-5, 5) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```
